---
title: "Phase and Syntactic Composition"
author: "Jonathan Brennan"
output:
  html_document:
    df_print: paged
    toc: yes
---

# Preliminaries

Analysis for

> Brennan, J. R., & Martin, A. E. (2019). Phase synchronization varies systematically with linguistic structure composition. Philosophical Transactions of the Royal Society B, 375. https://doi.org/10.1098/rstb.2019.0305

Originally run June 2019. Last tested August 2023 with 
R 4.2.0
brms_2.17.0 
emmeans_1.7.4 
bayestestR_0.13.1 
tidyverse_1.3.1

**Important Note**: 
Statistical output when the full pipeline is rerun is not numerically identical to published results. 
Differences appear to reflect small noise in how word word onset time-steps were aligned to EEG samples (e.g. rounding differences) and API changes in some key functions (e.g. `bayestestR`).
These differences do not appear to be consequential:
Data values correlates with original values at $r > 0.99$.
Statistical outputs are the same as original in terms of effect size and direction

Annotations: 

* CATEGORY: content words, function words
* BIN: : one, two, three+ closing brackets
* FREQ: delta (1-4), theta (4-8), gamma (30-50)

Analyses:

* **Power** at word onset
* **Inter-trial phase coherence** at word onset
* **Phase-to-phsae cross-frequency coupling**  at word onset
* **Phase-to power cross-frequency coupling** at word onset
* Compare each to **null distributions** by 
  - rotating the bracket-count measure +100 lags and recomputing
  - z-normalize all measures against these null distributions

Data are created with `run_power_phase_export_to_R.m` in MATLAB

```{r setup1, include=FALSE}

data_dir = '~/Desktop/alice-phase/'
setwd('/Users/jobrenn/Documents/Projects/alice-oscillations/ana5-ptrb-replicable')

```

```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning=FALSE)
options(digits=3)
library(tidyverse)
library(knitr)
library(gdata)
library(ggplot2)
library(circular)
library(reshape2)
library(brms)
library(broom)
library(emmeans)
library(bayestestR)
#library(loo)

options(mc.cores = 4) 
options(digits=3)

my_prior <- set_prior("student_t(3, 0, 3)", class="b")
```



## Load datasets

```{r load-data}
all <- read.csv('AliceChapterOne-Nodecounts.csv')
all <- subset(all, Trigger == 1)

# Make null bracket count
lagChange      <- 100
nWords         <- max(all$Order, na.rm=TRUE)
null.all       <- all
null.all$Order <- all$Order + lagChange
null.all$Order <- ifelse(null.all$Order > nWords, null.all$Order - nWords, null.all$Order)
null.all$null.brackets <- null.all$cfg.bu

all <- merge(all, null.all[c('Order', 'null.brackets')], by='Order')

# load data
datasets = dir(path=data_dir, pattern='*.csv')

for (i in seq_along(datasets)) {
  fname = paste(data_dir, datasets[i], sep='')
  if (i == 1) {
    hlb = read.csv(fname)
  } else {
    hlb = rbind(hlb, read.csv(fname))
  }
}

d <- merge(hlb, all, by.x='order', by.y='Order')

d$band <- factor(d$band, levels=c('delta', 'theta', 'gamma'))
d$bracket.count = d$cfg.bu
```

## Organize bracket annotations

generates Table 1 

```{r make.bins}
# make bracket BINS
d$bracket.bins <- ifelse(d$bracket.count == 1, 'one',   0)
d$bracket.bins <- ifelse(d$bracket.count == 2, 'two',   d$bracket.bins)
d$bracket.bins <- ifelse(d$bracket.count > 2, 'three+', d$bracket.bins)
d$bracket.bins <- factor(d$bracket.bins, levels=c('one', 'two', 'three+'))

# make null BINS
d$null.bins <- ifelse(d$null.brackets == 1, 'one',   0)
d$null.bins <- ifelse(d$null.brackets == 2, 'two',   d$null.bins)
d$null.bins <- ifelse(d$null.brackets > 2, 'three+', d$null.bins)
d$null.bins <- factor(d$null.bins, levels=c('one', 'two', 'three+'))

# set word category
d$category <- factor(d$islexical, labels=c('functional', 'lexical'))

# Summarize bin counts
print('Brackets per bin per subject, per category')
floor(table(d$bracket.bins, d$category) / 33 / 3)
#print('Brackets per bin per subject')
#floor(table(d$bracket.bins) / 33 / 3)

# Remove all NA rows (residual annotation mistakes)
orig_d <- d
d      <- subset(orig_d, !is.na(d$category))
d      <- subset(d, !is.na(d$bracket.bins))
d      <- subset(d, !is.na(d$null.bins))
1 - nrow(d) / nrow(orig_d) # 2.28% data-points removed

```

Save the data file that is published with the manuscript (Supplementary data file)

```{r}
d %>% 
  select(epochOrder = order, subject, band, onsetPower=ons_pwr, onsetPhase=ons_phs, word=AUDIOTEXT, segment, wordOnset = tmin, wordFrequency=logfreq, phraseCount=bracket.count, phraseBin=bracket.bins, phraseBinNull=null.bins, wordCategory=category) %>%
  write_csv('phase-by-bracket-data.csv')
```

## Content/Function length plots

Figure 1

```{r constituent.size}

sub = d[!duplicated(d$order),]

lex.length  = subset(sub, category=='lexical')$wordlength
func.length = subset(sub, category=='functional')$wordlength
word.length = sub$wordlength

par(mfrow=c(1,3))
hist(lex.length,     breaks=40, main='Content Words',    xlab='word   length (s)', bty='n')
hist(func.length,    breaks=40, main='Function Words', xlab='word   length (s)', bty='n', ylab='')
hist(word.length,    breaks=40, main='All Words', xlab='word   length (s)', bty='n', ylab='')
#hist(phrase.length,  breaks=40, main='Phrases',          xlab='phrase length (s)', bty='n', ylab='')

```


# Power

```{r power}

d %>%
  group_by(subject, bracket.bins, band, category) %>%
  summarize(ons_pwr = mean(ons_pwr)) ->
  brackets.pow

d %>%
  group_by(subject, null.bins, band, category) %>%
  summarize(ons_pwr = mean(ons_pwr)) -> 
  brackets.pow.null

brackets.pow$ons_pwr_null <- brackets.pow.null$ons_pwr

# z-score
brackets.pow %>%
  group_by(bracket.bins, band, category) %>%
  mutate(ons_pwr_z = (ons_pwr - mean(ons_pwr_null)) / sd(ons_pwr_null)) ->
  brackets.pow

```

## Power Plots

Figure 1A

```{r plot.power}

brackets.pow  %>%
  ggplot(aes(y = ons_pwr_z, x = band, col=factor(bracket.bins), fill=factor(bracket.bins))) +
  geom_point(position = position_jitterdodge(jitter.width=0.2, dodge.width=0.7), alpha=0.4, show.legend=FALSE) +
  geom_crossbar(col='black',stat='summary', fun.data=mean_se, width=0.5, position=position_dodge(width=0.7)) +
  facet_wrap( ~ category, nrow=1) +
  scale_y_continuous('z(Power)') +
  scale_x_discrete('') +
  scale_fill_manual(  name='Brackets: ', values=c('#d53e4f', '#f46d43', '#fdae61')) +
  scale_colour_manual(name='Brackets: ', values=c('#d53e4f', '#f46d43', '#fdae61')) +
  theme_minimal() +
  theme(legend.position= c(0.5, 0.9), legend.direction='horizontal') ->
  plot.brackets.pow

plot.brackets.pow

ggsave('figs/plot.brackets.pow.png', height=3, width=5)

```

## Power Stats

```{r pow.stats.brackets.aov}

aov.pow.brackets <- aov(ons_pwr_z ~ band * bracket.bins * category + 
                            Error(subject/(band * bracket.bins * category)), 
                          data=brackets.pow)
tidy(aov.pow.brackets)

```

Fit statistical model for power

```{r pow.stats.brackets.brm}
if (!file.exists('brm.pow.brackets.rData')) {
  brm.pow.brackets   <- brm(ons_pwr_z ~ band * bracket.bins * category + 
                           (1 + band * bracket.bins * category | subject), 
                         data = brackets.pow,
                         prior = my_prior, 
                         sample_prior = 'yes',
                         save_all_pars = TRUE,
                         verbose=FALSE)
  save(file='brm.pow.brackets.rData', 
       'brm.pow.brackets')
} else {
  load('brm.pow.brackets.rData')
}

```

Appendix A Table 2

```{r pow.stats.brackets.loo}
## marginal means and post-hoc comparisons
joint_tests(brm.pow.brackets)

# interaction of band:bracket.bins
brm.pow.brackets.emm <- pairs(
  emmeans(brm.pow.brackets, ~ band * bracket.bins * category), 
  by=c('band', 'category') 
  )

brm.pow.brackets.prior <- unupdate(brm.pow.brackets)
brm.pow.brackets.prior.emm <- pairs(
  emmeans(brm.pow.brackets.prior, ~ band * bracket.bins * category), 
  by=c('band', 'category') 
  )

brm.pow.brackets.bf <- describe_posterior(
  brm.pow.brackets.emm, bf_prior=brm.pow.brackets.prior.emm, 
  test='bf')

cbind(
  summary(rbind(brm.pow.brackets.emm)),
  BayesFactor=exp(brm.pow.brackets.bf$log_BF)
)

```

# Phase Consistancy: Intertrial Phase Coherence (ITPC)

Test for changes in phase alignment as combinatory demands change by examining **inter-trial phase clustering (ITPC)** per subject within each bin of bracket count.

ITPC comes from Eq. 19.1 in Cohen (2014) book "Analyzing Neural Time-Series Data":

$$ ITPC_{t,f} = \left| n^{-1} \sum\limits_{r=1}^n e^{ik} \right| $$ 
where $k$ is a phase angle in radians and $n$ is the total number of trials

```{r itpc}
# ITPC function takes a vector of phase angles
ITPC <- function(ph) { abs(mean(exp(1i * ph ) ) ) }

# group by bins
d %>%
  filter(!is.na(bracket.bins)) %>%
  group_by(subject, bracket.bins, band, category) %>%
  summarize(ons_itpc = ITPC(as.numeric(ons_phs))) ->
  brackets.itpc

d %>%
  filter(!is.na(null.bins)) %>%
  filter(!is.na(category)) %>%
  group_by(subject, null.bins, band, category) %>%
  summarize(ons_itpc = ITPC(as.numeric(ons_phs))) ->
  brackets.itpc.null

brackets.itpc$ons_itpc_null <- brackets.itpc.null$ons_itpc

# z-score
brackets.itpc %>%
  group_by(bracket.bins, band, category) %>%
  mutate(ons_itpc_z = (ons_itpc - mean(ons_itpc_null)) / sd(ons_itpc_null)) ->
  brackets.itpc

```

## ITPC Plots

Figure 1B

```{r plot.itpc}

brackets.itpc  %>%
  ggplot(aes(y = ons_itpc_z, x=band, col=factor(bracket.bins), fill=factor(bracket.bins))) +
  geom_point(alpha=0.4, position=position_jitterdodge(jitter.width=0.2, dodge.width=0.7), show.legend=FALSE) +
  geom_crossbar(col='black', stat='summary', fun.data=mean_se, width=0.5, position=position_dodge(width=0.7)) +
  facet_wrap( ~ category, nrow=1) +
  scale_y_continuous('z(ITPC)') +
  scale_x_discrete('') +
  scale_fill_manual(  name='Brackets: ', values=c('#d53e4f', '#f46d43', '#fdae61')) +
  scale_colour_manual(name='Brackets: ', values=c('#d53e4f', '#f46d43', '#fdae61')) +
  theme_minimal() +
  theme(legend.position='none') ->
  plot.bracket.itpc

plot.bracket.itpc

ggsave('figs/plot.bracket.itpc.png', height=3, width=5)

```

## ITPC Stats

```{r itpc.stats.brackets.aov}
aov.itpc.brackets <- aov(ons_itpc_z ~ band * bracket.bins * category + 
                            Error(subject/(band * bracket.bins * category)),
                         data = brackets.itpc)
tidy(aov.itpc.brackets)
```

Fit ITPC stats

```{r itpc.stats.brackets.brm}
# must set sample_prior = 'yes' to test *point hypotheses* with hypothesis()

if (!file.exists('brm.itpc.brackets.rData')) {
  brm.itpc.brackets   <- brm(ons_itpc_z ~ band * bracket.bins * category + 
                           (1 + band * bracket.bins * category | subject), 
                         data = brackets.itpc,
                         sample_prior = 'yes',
                         save_all_pars = TRUE,
                         prior = my_prior,
                         verbose=FALSE)
  save(file='brm.itpc.brackets.rData', 
       'brm.itpc.brackets')
} else {
  load('brm.itpc.brackets.rData')
}

```
Appendix A Table 3

```{r itpc.stats.brackets.loo}
## marginal means and post-hoc comparisons
joint_tests(brm.itpc.brackets)

brm.itpc.brackets.emm <- pairs(
  emmeans(brm.itpc.brackets, ~ band * bracket.bins * category), 
  by=c('band', 'category') )

brm.itpc.brackets.prior <- unupdate(brm.itpc.brackets)
brm.itpc.brackets.prior.emm <- pairs(
  emmeans(brm.itpc.brackets.prior, ~ band * bracket.bins * category), 
  by=c('band', 'category') 
  )

brm.itpc.brackets.bf <- describe_posterior(
  brm.itpc.brackets.emm, bf_prior=brm.itpc.brackets.prior.emm, 
  test='bf')

cbind(
  summary(rbind(brm.itpc.brackets.emm)),
  BayesFactor=exp(brm.itpc.brackets.bf$log_BF)
)

```

## Phase histograms

Figure 1C

```{r circular.plots}

ggplot(d, aes(x=ons_phs, col=bracket.bins)) +  
  geom_histogram(aes(y=..density..), bins=30, fill='white') +
  coord_polar(clip='off') +
  facet_grid(bracket.bins~category+band) +
  scale_colour_manual(name='Brackets: ', values=c('#d53e4f', '#f46d43', '#fdae61')) +
  scale_x_continuous('', limits=c(-pi, pi), breaks = c(0, pi), labels=c(0, expression(pi))) +
  theme_minimal() +
  theme(legend.position='none') 

ggsave('figs/phase.histograms.png', height=3, width=5)


```

Supplemental Figure 1

```{r phase.by.subj, fig.height=8, fig.width=7}
# Single-subject phase density plots, just on functional words in the delta-band 

ggplot(subset(d, category == 'functional' & band == 'delta') , aes(x=ons_phs, fill=bracket.bins)) +  
  geom_density(alpha=0.4) +
  facet_wrap(~subject) +
  scale_x_continuous('', limits=c(-pi, pi), breaks = c(0, pi), labels=c(0, expression(pi))) +
  theme_classic() +
  theme(legend.position=c(0.75, 0.05), legend.direction='horizontal') +
  ggtitle('Delta-band phase density plots, function words')


ggsave('figs/phase.single.subject.png', height=8, width=7)
# 

```

# Cross frequency coupling: Power-Power correlation (P-P)

```{r cfc.pow.cor}
nData    <- length(unique(d$subject)) * 3 * 3 * 2 # 3 bracket bins, 3 band.pairs, 2 categories
fisher.z <- function(r) { 0.5 * (log(1+r) - log(1-r)) }

cfc.pwr.brackets <- as_tibble(matrix(nrow=nData, ncol=6), .name_repair='minimal')  
names(cfc.pwr.brackets) <- c('subject', 'category', 'bracket.bins', 'band.pair', 'z.r', 'null.z.r')

i = 1
for (s in unique(d$subject)) {
  for (c in unique(d$category)) {
    if (is.na(c)) { next }
    for (b in unique(d$bracket.bins)) {
      if (is.na(b)) { next }
      f.track = c() # track lower frequency
      for (f1 in unique(d$band)) {
        f.track = c(f.track, f1)
        for (f2 in unique(d$band)) { # order of band ensures f1 <= f2
          if (f2 %in% f.track) { next }
          first  = subset(d, subject == s & category == c & bracket.bins == b & band == f1)$ons_pwr
          second = subset(d, subject == s & category == c & bracket.bins == b & band == f2)$ons_pwr
          cfc.pwr.brackets$z.r[i]           <- fisher.z(cor(first, second))
          
          first  = subset(d, subject == s & category == c & null.bins == b & band == f1)$ons_pwr
          second = subset(d, subject == s & category == c & null.bins == b & band == f2)$ons_pwr
          cfc.pwr.brackets$null.z.r[i]      <- fisher.z(cor(first, second))
          
          cfc.pwr.brackets$subject[i]       <- s
          cfc.pwr.brackets$category[i]      <- c
          cfc.pwr.brackets$bracket.bins[i]  <- b
          cfc.pwr.brackets$band.pair[i]     <- paste(f1, f2, sep='.')
          i <- i + 1
        }
      }
    }
  }
}

cfc.pwr.brackets$bracket.bins <- factor(cfc.pwr.brackets$bracket.bins, levels=c('one', 'two', 'three+'))

# add z-score against null values
cfc.pwr.brackets %>%
  group_by(category, bracket.bins, band.pair) %>%
  mutate(z.z.r = (z.r - mean(null.z.r)) / sd(null.z.r)) ->
  cfc.pwr.brackets

```

## P-P Plots

Figure 1D

```{r plot.cfc.pow}

cfc.pwr.brackets  %>%
  ggplot(aes(y = z.z.r, x=band.pair, col=factor(bracket.bins), fill=factor(bracket.bins))) +
  geom_point(alpha=0.4, position=position_jitterdodge(jitter.width=0.2, dodge.width=0.7), show.legend=FALSE) +
  geom_crossbar(col='black', stat='summary', fun.data=mean_se, width=0.5, position=position_dodge(width=0.7)) +
  facet_wrap( ~ category, nrow=1) +
  scale_y_continuous('z(r), power') +
  scale_x_discrete('') +
  scale_fill_manual(  name='Brackets: ', values=c('#d53e4f', '#f46d43', '#fdae61')) +
  scale_colour_manual(name='Brackets: ', values=c('#d53e4f', '#f46d43', '#fdae61')) +
  theme_minimal() +
  theme(legend.position= 'none') ->
  plot.cfc.pwr.bra

plot.cfc.pwr.bra

ggsave('figs/plot.cfc.pwr.bra.png', height=3, width=5)

```

## P-P Stats

```{r cfc.pow.cor.brackets.stat.aov}

aov.cfc.pwr.bracket <-  aov(z.z.r ~ band.pair * bracket.bins * category + 
                               Error(subject / (band.pair * bracket.bins * category)), 
                         data = cfc.pwr.brackets)

tidy(aov.cfc.pwr.bracket)
```

Fit P-P stats

```{r cfc.pow.cor.brackets.stat.brm}

if (!file.exists('brm.cfc.pwr.bracket.rData')) {
  brm.cfc.pwr.brackets  <- brm(z.z.r ~ band.pair * bracket.bins * category + 
                           (1 + band.pair * bracket.bins * category | subject), 
                         data = cfc.pwr.brackets,
                         sample_prior = 'yes',
                         save_all_pars = TRUE,
                         prior = my_prior, 
                         verbose=FALSE)
  save(file='brm.cfc.pwr.bracket.rData', 'brm.cfc.pwr.brackets')
} else {
  load('brm.cfc.pwr.bracket.rData')
}

```

Appendix A Table 4

```{r cfc.pow.cor.brackets.stat.loo}
## marginal means and post-hoc comparisons
joint_tests(brm.cfc.pwr.brackets)

brm.cfc.pwr.brackets.emm <- pairs(
  emmeans(brm.cfc.pwr.brackets, ~ band.pair * bracket.bins * category), 
  by=c('band.pair', 'category') )

brm.cfc.pwr.brackets.prior <- unupdate(brm.cfc.pwr.brackets)
brm.cfc.pwr.brackets.prior.emm <- pairs(
  emmeans(brm.cfc.pwr.brackets.prior, ~ band.pair * bracket.bins * category), 
  by=c('band.pair', 'category') 
  )

brm.cfc.pwr.brackets.bf <- describe_posterior(
  brm.cfc.pwr.brackets.emm, bf_prior=brm.cfc.pwr.brackets.prior.emm, 
  test='bf')

cbind(
  summary(rbind(brm.cfc.pwr.brackets.emm)),
  BayesFactor=exp(brm.cfc.pwr.brackets.bf$log_BF)
)


```

# Cross-frequency coupling: Phase-amplitude coupling (PAC)

From Cohen 2014 ch. 30 p. 413:

$$PAC = \left| n^{-1} \sum\limits_{t=1}^n a_te^{i\phi_t} \right|$$

Where $t$ is the time-point (here, onset), $a$ is power at the higher frequency, $\phi$ is the phase angle (radians) of the lower frequency and $n$ is the total number of time-points.

Following Cohen, use raw power, not transformed, so PAC is positive and test with a non-parametric statistic (shift the phase/amplitude time indices to create null distribution)

```{r cfc.pac}
# exp(1) gives e
# pw is power at the *higher frequency*, ph is phase at the *lower frequency*
pac <- function(pw, ph) { abs(mean(pw * exp(1)^(1i * ph) ) ) }

brackets.pac <- as_tibble(matrix(nrow=nData, ncol=6))
names(brackets.pac) <- c('subject', 'category', 'bracket.bins', 'band.pair', 'pac', 'null.pac')
i = 1
for (s in unique(d$subject)) { # for each subject, each BIN, and each unique band.pair...
  for (c in unique(d$category)) {
    if (is.na(c)) { next }
    for (b in unique(d$bracket.bins)) {
      if (is.na(b)) { next }
      track.f = c() # keep track of lower frequencies
      for (f1 in unique(d$band)) {
        track.f = c(track.f, f1)
        for (f2 in unique(d$band)) { # factor order ensures f1 <= f2
          if (f2 %in% track.f) { next }
          # ... get vector PH of phases from the lower frequency
          ph <- subset(d, subject == s & bracket.bins == b & band == f1 & category == c)$ons_phs
          # ... get vector PW of powers from the higher frequency
          pw <- subset(d, subject == s & bracket.bins == b & band == f2 & category == c)$ons_pwr
          # ... return pac(P,H)
          brackets.pac[i,'pac']          <- pac(pw, as.numeric(ph))

          # get null effects
          ph <- subset(d, subject == s & null.bins == b & band == f1 & category == c)$ons_phs
          pw <- subset(d, subject == s & null.bins == b & band == f2 & category == c)$ons_pwr
          brackets.pac[i,'null.pac']     <- pac(pw, as.numeric(ph))
          brackets.pac[i,'subject']      <- s
          brackets.pac[i,'category']     <- c
          brackets.pac[i,'bracket.bins'] <- b
          brackets.pac[i,'band.pair']    <- paste(f1, f2, sep='.')
          i = i + 1
        }
      }
    }
  }
}

brackets.pac$bracket.bins <- factor(brackets.pac$bracket.bins, levels=c('one', 'two', 'three+'))

# add z-score against null values
brackets.pac %>%
  group_by(category, bracket.bins, band.pair) %>%
  mutate(pac_z = (pac - mean(null.pac)) / sd(null.pac)) ->
  brackets.pac


```

## PAC Plots

Figure 2E

```{r plot.cfc.pac}

brackets.pac  %>%
  ggplot(aes(y = pac_z, x = band.pair, col=factor(bracket.bins), fill=factor(bracket.bins))) +
  geom_point(alpha=0.4, position=position_jitterdodge(jitter.width=0.2, dodge.width=0.7)) +
  geom_crossbar(col='black', stat='summary', fun.data=mean_se, width=0.5, position=position_dodge(width=0.7)) +
  facet_wrap( ~ category, nrow=1) +
  scale_y_continuous('z(PAC)') +
  scale_x_discrete('') +
  scale_fill_manual(  name='Brackets: ', values=c('#d53e4f', '#f46d43', '#fdae61')) +
  scale_colour_manual(name='Brackets: ', values=c('#d53e4f', '#f46d43', '#fdae61')) +
  theme_minimal() +
  theme(legend.position= 'none') ->
  plot.pac.brackets

plot.pac.brackets

ggsave('figs/plot.bracket.pac.png', height=3, width=5)

```

## PAC Stats

```{r cfc.pac.brackets.stat}

aov.pac.brackets <- aov(pac_z ~ bracket.bins * band.pair * category + 
                          Error(subject/(bracket.bins * band.pair * category)), 
                       data=brackets.pac)
tidy(aov.pac.brackets)
```

Fit PAC stats

```{r cfc.pac.brackets.stat.brm}

if (!file.exists('brm.pac.brackets.rData')) {
  brm.pac.brackets  <- brm(pac_z ~ band.pair * bracket.bins * category + 
                           (1 + band.pair * bracket.bins * category | subject), 
                         data = brackets.pac,
                         sample_prior = 'yes',
                         save_all_pars = TRUE,
                         prior = my_prior,
                         verbose=FALSE)
  save(file='brm.pac.brackets.rData', 
       'brm.pac.brackets')
} else {
  load('brm.pac.brackets.rData')
}

```


Appendix A Table 5

```{r cfc.pac.brackets.stat.loo}
joint_tests(brm.pac.brackets)

brm.pac.brackets.emm <- pairs(
  emmeans(brm.pac.brackets, ~ band.pair * bracket.bins * category), 
  by=c('band.pair', 'category') )

brm.pac.brackets.prior <- unupdate(brm.pac.brackets)
brm.pac.brackets.prior.emm <- pairs(
  emmeans(brm.pac.brackets.prior, ~ band.pair * bracket.bins * category), 
  by=c('band.pair', 'category') 
  )

brm.pac.brackets.bf <- describe_posterior(
  brm.pac.brackets.emm, bf_prior=brm.pac.brackets.prior.emm, 
  test='bf')

cbind(
  summary(rbind(brm.pac.brackets.emm)),
  BayesFactor=exp(brm.pac.brackets.bf$log_BF)
)


```


